The demographic costs of sexually antagonistic selection in partially selfing populations

### 2021-12-10 doc/ResponseLetter-R1.tex

In addition to the clarifications/edits and other points raised by the reviewers, I include a handful of minor edits/typos here of my own:

\noindent (1) Abstract, perhaps add a very general and short explanation of the term sexually antagonistic polymorphism.
\begin{quote}
	{\itshape ...}
\end{quote}

\noindent (2) Line 165; "fertility"
\begin{quote}
	{\itshape Fixed.}
\end{quote}

\noindent (3) Table 1, definition of inbreeding depression terms; it would be clearer to define each of the 4 terms separately, "Inbreeding depression terms for ovule viability, juvenile survival, adult survival, and transition rates respectively"
\begin{quote}
	{\itshape Fixed}
\end{quote}

\noindent (4) \textcolor{red}{Figure 1; I generally prefer axes labels to have clear meanings, rather than just symbols – here, the vertical and horizontal axes give the female (vertical) vs. male (horizontal) selection on reproductive function.}
\begin{quote}
	{\itshape ...}
\end{quote}

\noindent (5) Lines 305, 306, and 478; the use of the term "fixates" rather than "fixes" is not standard, I would change to "fixes" in all cases.
\begin{quote}
	{\itshape Fixed \textcolor{red}{Note to Colin: I find that fixes sounds a bit odd? I changed it to "goes to fixation" in one instance. I trust your final judgement on this as our native.}}
\end{quote}

\noindent (6) \textcolor{red}{Figure 4; same comment as for Figure 1.}
\begin{quote}
	{\itshape ...}
\end{quote}

@@

\noindent Apart from that, I have only minor comments, that I list below (roughly in decreasing order of importance):
\bigskip

\noindent -- Model: lines 114-195
\begin{itemize}
	\item The matrices H are mentioned before they are introduced. (Therefore, the parenthesis in lines 123-124 cannot be understood by the reader). And then are only defined in the following section	
	\begin{itemize}
		\item It would make sense to reorder the model presentation by presenting the building blocks (arguably U, F and H matrices, but I'll come back to H in the next point), what they do, what they are for; then provide some properties (F independent from selfing/crossing, H from stage, etc ...) and then go into the details of their construction
	\end{itemize}
	\item The main issue I have here, is that H is introduced, but then disappears from the output (see equations 12 and 13 for instance that use H but it does not appear, or the general output eq 9)
	\begin{itemize}
		\item This suggests combining F and H directly into an object.
	\end{itemize}
\end{itemize}

\begin{quote}
	{\itshape ... our response}
\end{quote}

\begin{itemize}
	\item More generally, the matrix formulations of the model may be a bit hard to read for the general reader of AmNat. Getting rid of H (or to the contrary, using it all the way, thus simplifying eq 12 and 13 for instance) and not giving a name to matrix W (which is just a bunch of 0s 1s and 1/2s) are steps that can make it easier for the reader to follow.
	\item It is quite confusing that HX is written as a constant (eq 7) whilst it depends on the frequencies and that HS is written as dependent (eq6) whilst it is constant
	\begin{quote}
	{\itshape That's a mistake on our part, apologies.}
\end{quote}
	\item As the building blocks are given expressions (eq 10 and 11 for instance) it would be nice to mention the fact that the elements in these will be characterize later (otherwise the reader thinks she/he should already know what the expression of, say, UsAA is)
\end{itemize}

@@

\noindent-- Fig 4 : they are supposed to be dashed lines and dotted lines but all I see are dashed lines with various dashes. I found 4B in particular, hard to understand at first sight.

\begin{quote}
	{\itshape ... our response}
\end{quote}
\bigskip

\noindent Typos:
\smallskip

\noindent-- The 0s representing 0 matrices in eq 5 are not bold
\begin{quote}
	{\itshape Fixed}
\end{quote}

\noindent-- In same equation 1w is not defined
\begin{quote}
	{\itshape Fixed}
\end{quote}

\noindent-- Line 142, 2 "the"s
\begin{quote}
	{\itshape Fixed}
\end{quote}

### 2021-12-10 doc/SA-Hermaphrodites-wDemography-R1.tex
Reviews1_part1
Edit methods,  paragraph intro about models combined population structure&genetics, fixed some minor things from the minor edits list (see response letter)

Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant (and often infinite) size, and therefore do not consider the population dynamical consequences of evolutionary change. In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \citealt {johnston2013life, merot2020balancing}). To take into account the consequences of population dynamics under SA selection, a model is required that links changes in genotype distributions to population dynamics. The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then have explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}). \citet{deVriesCaswell2019a} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.

@@

Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of partially-selfing simultaneous hermaphrodites. The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}. In fact, the model reduces to the stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing. A full derivation of the model and analyses is presented in the Online Supplementary Material, and all computer code necessary to reproduce the results are available at \url{GitHub Link omitted for anonymity}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.

@@

The population projection matrix $\tilde{\mbf{A}}$ is constructed from four sets of matrices representing the demographic and genetic processes: The matrices $\mbf{U}^{S}_{i}$ and $\mbf{U}^{X}_{i}$ contain transition and survival probabilities for each genotype, produced by selfing and outcrossing respectively. The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contribution of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation. We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime}_{i}$. Deviations from this assumption are straightforward to incorporate but beyond the scope of this paper.

For the purpose of this article, we assume that mating is random with respect to stage and hence that the parent-offspring map is the same for all stages (i.e., $\mbf{H}^S_{j}(\tilde{\mbf{n}}) = \mbf{H}^S(\tilde{\mbf{n}})$, and $\mbf{H}^X_{j}(\tilde{\mbf{n}}) = \mbf{H}^X(\tilde{\mbf{n}})$; where $j \in \{1,\omega\}$). The matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X(\tilde{\mbf{n}})$ contain the population genetic processes and are presented in the next section.

%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Mating and offspring production under partial selfing}
 Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}. That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined. For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}. Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion.

Because we assume that inbreeding depression only affects offspring viability, the allele frequency in the male mating population is obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
\begin{linenomath*}
\begin{equation}
 	\mbf{p}^{\prime}=\frac{(\mbf{n}^X + \mbf{n}^S)}{\| (\mbf{n}^X + \mbf{n}^S)\|}.
 \end{equation}
\end{linenomath*}
Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the following equation
\begin{linenomath*}
\begin{equation}
		\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}= 
		\left(
			\begin{array}{ccc}
				\mbf{1}^{\intercal}_{\omega} & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & 0 \\
				0 & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{1}^{\intercal}_{\omega} \\
			\end{array} \right)
		\left(
			\begin{array}{ccc}
				\mbf{F}^{\prime}_{AA} & 0 & 0 \\
				0 & \mbf{F}^{\prime}_{Aa} & 0 \\
				0 & 0 & \mbf{F}^{\prime}_{aa} \\
			\end{array} \right)
		\left(
			\begin{array}{c}
				\mbf{p}^{\prime}_{AA} \\
				\mbf{p}^{\prime}_{Aa} \\
				\mbf{p}^{\prime}_{aa} \\
			\end{array} \right).
\end{equation}
\end{linenomath*}

\noindent The matrix $\mbb{F}^{\prime}$ is composed of genotype-specific fertility matrices, and operates on the the vector of genotype frequencies to give their relative contribution of each genotype to the gamete pool. The matrix $\mbf{W}^{\prime}$ converts these contributions to allele numbers. Normalizing the resulting vector gives the allele frequencies in the male gamete pool:
\begin{linenomath*}
\begin{equation} \label{eq:maleGametePool}
	\left(
		\begin{array}{c}
			q^{\prime}_{A} \\
			q^{\prime}_{a} \\
		\end{array} \right) = 
			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime} \|} = 
				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)\|}.
\end{equation}
\end{linenomath*}
The key difference between selfing and outcrossing reproduction can be seen in the parent-offspring maps, which reflect the joint effects of meiosis and mating on the distribution of offspring genotypes.
% \textcolor{red}{
% I'm doubting whether to move the next paragraph to the appendices as well.  Equation (\ref{eq:maleGametePool}) is all we strictly need. If we keep the parent-offspring maps, we can explain in the next section how the matrices $\mcal{F}^S$ and $\mcal{F}^X$ have the same structure. Would that be helpful? Or would it be easier to just explain in the next section what the structure of $\mcal{F}^S$ and $\mcal{F}^X$ is? }

The parent-offspring map for outcrossing is a function of the allele frequencies in the male gamete pool, and is given by 
\begin{linenomath*}
\begin{equation} \label{eq:HX}
	\mbf{H}^X = 
			\left(
			\begin{array}{ccc}
				q^{\prime}_{A} & \frac{1}{2} q^{\prime}_{A} & 0 \\
				q^{\prime}_{a} & \frac{1}{2} & q^{\prime}_{A}  \\
				0 & \frac{1}{2} q^{\prime}_{a} & q^{\prime}_{a} \\
			\end{array} \right).
\end{equation}
\end{linenomath*}
\noindent From left to right, the columns of matrix $\mbf{H}^X$ give the genotype distribution of outcrossed offspring produced by a maternal parent of each genotype ($AA$, $Aa$, and $aa$ respectively).

The parent-offspring map for reproduction via self-fertilization differs from the outcrossing parent-offspring map. As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection via the male sex function from external factors (e.g., pollinator visitation) relative to outcrossing (\citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}). Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during and after meiosis: 
\begin{linenomath*}
\begin{equation} \label{eq:HS}
	\mbf{H}^S(\tilde{\mbf{n}}) = 
			\left(
			\begin{array}{ccc}
				1 & 1/4 & 0 \\
				0 & 1/2 & 0 \\
				0 & 1/4 & 1 \\
			\end{array} \right).
\end{equation}
\end{linenomath*}
\noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively. As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertilitiy matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS}) and Eq(\ref{eq:BlkFX}) below).

Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability. In the Online Supplement, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).

%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Population projection}
Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$, and the parent-offspring matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X$), we construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}, see Online Supplementary Material for the step by step construction of the model. The matrix that projects the eco-evolutionary dynamics is:
\begin{linenomath*}
\begin{equation} \label{eq:Atilde}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
			\end{array} \right)}_{\tilde{\mbf{F}}},
\end{equation}
\end{linenomath*}
where $C$ denotes the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and  $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.

The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of extant individuals produced by selfing or outcrossing ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$). The survival matrices for individuals produced through selfing and outcrossing are,
\begin{linenomath*}
\begin{eqnarray} 
	\mcal{U}^S  = 
		\left(
			\begin{array}{ccc}
				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
			\end{array} \right),\label{eq:BlkUS}\\
				\mcal{U}^X  = 
		\left(
			\begin{array}{ccc}
				\mbf{U}^{X}_{AA} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{U}^{X}_{Aa} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{U}^{X}_{aa} \\
			\end{array} \right).\label{eq:BlkUX}
\end{eqnarray}
\end{linenomath*}
Since individuals cannot change their genotype once they are born, the survival matrices are block diagonal.  Similarly, we construct fertility matrices for individuals produced through selfing and outcrossing, 
\begin{linenomath*}
\begin{equation} \label{eq:BlkFS}
	\mcal{F}^S = 
			\left(
			\begin{array}{ccc}
				\mbf{F}_{AA} & \frac{1}{4} \mbf{F}_{Aa} & 0 \\
				0 & \frac{1}{2} \mbf{F}_{Aa} & 0 \\
				0 & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
			\end{array} \right), 
\end{equation}
\end{linenomath*}
\noindent and 
\begin{linenomath*}
\begin{equation} \label{eq:BlkFX}
	\mcal{F}^X = 
			\left(
			\begin{array}{ccc}
				q^{\prime}_{A} \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} & 0 \\ 
				q^{\prime}_{a} \mbf{F}_{AA} & \frac{1}{2} \mbf{F}_{Aa} & q^{\prime}_{A} \mbf{F}_{aa}  \\ 
				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} & q^{\prime}_{a} \mbf{F}_{aa}  \\
			\end{array} \right),
\end{equation}
\end{linenomath*}
\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}).

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[htbp]
\centering
\caption{\bf Definition of terms.}
\label{tab:Terms}
\begin{tabular}{ l p{0.7\linewidth} r }
 \toprule
Symbol & Definition & Dimension \\
\hline
$g$      & Number of genotypes ($3$; $AA$, $Aa$, and $aa$) & \\
$\omega$ & Number of stages ($2$; juvenile and adult) & \\
$\tilde{\mbf{n}}$ & Joint stage $\times$ genotype vector & $2 \omega g \times 1$ \\
$\tilde{\mbf{p}}$ & Joint stage $\times$ genotype frequency vector & $2 \omega g \times 1$ \\
$\mbf{U}^{S}_{i}$,\, $\mbf{U}^{X}_{i}$ & Genotype-specific transition and survival matrices & $\omega \times \omega$ \\
$\mcal{U}^{S},\, \mcal{U}^{X}$ & Block diagonal selfed/outcrossed survival matrices & $\omega g \times \omega$ g\\
$\mbf{F}_{i}$,\, $\mbf{F}^{\prime}_{i}$ & Genotype-specific fertility matrices & $\omega \times \omega$ \\
$\mcal{F}^{S},\, \mcal{F}^{X}$ & Block diagonal selfed/outcrossed fertility matrix & $\omega g \times \omega$g \\
$\mbf{H}^{S}_{i}$,\, $\mbf{H}^{X}_{i}$ & Parent-offspring genotype maps & $g \times g$ \\
$\mbf{W}$   & Allele segregation matrix & $2 \times \omega $g \\
% $\mbf{K}$   & Vec-permutation matrix & $g \times g$ \\
% $\mbf{I}$   & Identity matrix & given in subscript \\
$q^{\prime}_{A}$,\, $q^{\prime}_{a}$ & Allele frequencies in male gamete pool & $1$ \\
$\tilde{\mbf{A}}(\tilde{\mbf{p}})$ & Population projection matrix & $2 \omega g \times 2\omega g$ \\

$s_f,\,s_m$ & Selection coefficients through female and male reproductive function & \\
$h_f,\,h_m$ & Dominance coefficients through female and male reproductive function & \\
$\sigma_j,\, \sigma_a$ & Survival rates for juvenile and adult stages & \\
$\gamma$ & Transition rate from juvenile to adult stages \\
$C$ & The population rate of self-fertilization & \\
$\delta,\, \delta_j,\,\delta_a,\,\delta_{\gamma}$ & Inbreeding depression terms for ovule viability, survival, and transition rates respectively & \\
$f_i,\,f^{\prime}_i$ & Adult fertilities through female and male sex functions respectively & \\

%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}

We now construct and analyze a genotype $\times$ stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions. For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active. Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect. Following convention, we parameterize the fertility component of fitness for each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).

@@

The fertility matrices through female and male function are
\begin{linenomath*}
\begin{equation}\label{eq:FS}
	\mbf{F}_{i} = \left(
					\begin{array}{cc}
						0 & f w_{i} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\end{linenomath*}
\noindent and
\begin{linenomath*}
\begin{equation}\label{eq:FX}
	\mbf{F}^{\prime}_{i} = \left(
					\begin{array}{cc}
						0 & f^{\prime} w^{\prime}_{i} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\end{linenomath*}

\noindent where $f$ and $f^{\prime}$ represent adult fertilities, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).

@@

>>>>

Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant (and often infinite) size, and therefore do not consider the population dynamical consequences of evolutionary change. In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \citealt {johnston2013life, merot2020balancing}). To take into account the consequences of population dynamics under SA selection, a model is required that links changes in genotype distributions to population dynamics. \textcolor{blue}{One of the earliest such models \citep{roughgarden1971density} combined a population genetic model with a density-dependent population model with no intraspecific demographic structure. Charlesworth then combined an age-structured population model with population genetics in a series of papers, consolidated in \cite{BCharlesworth1991}. \cite{orive1995senescence} extended this work to study life-history evolution in organisms with both clonal and sexual reproduction (see also \cite{orive2001somatic}).  \cite{ronce2000kin} added spatial structure and stochasticity to age-structured genetic models to model kin selection, a body of work consolidated in \cite{rousset2004genetic}. Of course there are many other approaches to studying eco-evolutionary dynamics, e.g. combining quantitative genetics models with demographic structure (e.g. \cite{coulson2006putting,barfield2011evolution,childs2016evolution,orive2017effects}) or adaptive dynamics approaches (e.g. \cite{dieckmann2002adaptive,durinx2008adaptive,ten2020habitat}). For a comprehensive review of the eco-evolutionary literature, see \cite{fussmann2007eco}, or more recently, \cite{govaert2019eco}.}

The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then have explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}). \citet{deVriesCaswell2019a} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.

@@

Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of partially-selfing simultaneous hermaphrodites. \textcolor{blue}{Rather than go through the full mathematical derivation of the general model, we showcase a two-stage example of our model and discuss the biological assumptions that go into the model. The derivation of the general model for any number of age or stage classes is in the Online Supplementary Material.} The derivation follows closely the logic presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}. In fact, the model reduces to the stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing. All computer code necessary to reproduce the results are available at \url{https://anonymous.4open.science/r/SA-Hermaphrodites-wDemography-BF66}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.

@@

The population projection matrix $\tilde{\mbf{A}}$ is constructed from four sets of matrices representing the demographic and genetic processes: The matrices $\mbf{U}^{S}_{i}$ and $\mbf{U}^{X}_{i}$ contain transition and survival probabilities for each genotype, produced by selfing and outcrossing respectively. The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contribution of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation. We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime}_{i}$, respectively. Deviations from this assumption are straightforward to incorporate but beyond the scope of this paper.

\subsubsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}

\textcolor{blue}{To demonstrate how to construct and analyze a genotype $\times$ stage-classified model, we will now construct an example model for a hypothetical species with intralocus sexual conflict via the two sex functions. }For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active. Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect. Following convention, we parameterize the fertility component of fitness for each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).

@@

The fertility matrices through female and male function are
\begin{linenomath*}
\begin{equation}\label{eq:F_female}
	\mbf{F}_{i} = \left(
					\begin{array}{cc}
						0 & f w_{i} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\end{linenomath*}
\noindent and
\begin{linenomath*}
\begin{equation}\label{eq:F_male}
	\mbf{F}^{\prime}_{i} = \left(
					\begin{array}{cc}
						0 & f^{\prime} w^{\prime}_{i} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\end{linenomath*}

\noindent where $f$ and $f^{\prime}$ represent maximum adult fertilities, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).

@@

\subsubsection*{Mating and offspring production under partial selfing}
 Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}. That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined. For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}. Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion.

\textcolor{blue}{ For the purpose of this article, we also assume that females in the population all have the same mating preferences regardless of female age, stage, or genotype. The model can be extended to consider assortative mating, but that is beyond the scope of the current paper. In addition, we assume that males produced through selfing or through outcrossing do not differ in their mating success.  Given these assumptions, the allele frequencies in the male mating population can obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
\begin{linenomath*}
\begin{equation}
 	\mbf{p}^{\prime}=\frac{(\mbf{n}^X + \mbf{n}^S)}{\| (\mbf{n}^X + \mbf{n}^S)\|}.
 \end{equation}
\end{linenomath*}}

\textcolor{blue}{Males contribute  gametes to the population gamete pool according to their mating success, which depends on their stage and genotype at the SA locus. The allele frequencies in the male gamete pool are then calculated as follows (see Online Supplementary Materials for a derivation of this equation),
\begin{linenomath*}
\begin{equation} \label{eq:maleGametePool}
	\left(
		\begin{array}{c}
			q^{\prime}_{A} \\
			q^{\prime}_{a} \\
		\end{array} \right) \propto 
				\left(
		\begin{array}{c}
			\mbf{1}^{\intercal}_{\omega}\mbf{F}^{\prime}_{AA}\mbf{p}^{\prime}_{AA}+ \frac{1}{2} \mbf{1}^{\intercal}_{\omega}  \mbf{F}^{\prime}_{Aa} \mbf{p}^{\prime}_{Aa}  \\
		\frac{1}{2} \mbf{1}^{\intercal}_{\omega}  \mbf{F}^{\prime}_{Aa} \mbf{p}^{\prime}_{Aa} + \mbf{1}^{\intercal}_{\omega}\mbf{F}^{\prime}_{aa}\mbf{p}^{\prime}_{aa}\\
		\end{array} \right),
\end{equation}
\end{linenomath*}
where equality is obtained by normalizing the frequencies such that they add to one. The first entry is simply summing the stage-specific contributions to the gamete pool of heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions to the gamete pool of heterozygotes and $aa$ homozygotes. }

\textcolor{blue}{ FWe now have all the ingredients to calculate the allele frequencies in the (male) gamete pool for the two stage example model in equations \ref{eq:F_male}, \ref{eq:maleGametePool}, and Table \ref{tab:Fitness}. Denoting the proportion of adults of genotype $i$ with $p^A_i$, the allele frequencies in the male gamete pool are given by 
\begin{linenomath*}
\begin{equation} \label{eq:maleGametePool_example}
\left(
		\begin{array}{c}
			q^{\prime}_{A} \\
			q^{\prime}_{a} \\
		\end{array} \right) = 
				\frac{1}{(1-s_m)p^A_{aa}+ (1-hs_m)p^A_{Aa}+p_{aa}}\left(
		\begin{array}{c}
(1-s_m)p^A_{AA}	+\frac{1}{2}(1-hs_m)p^A_{Aa}\\
p^A_{aa}+\frac{1}{2}(1-hs_m)p^A_{Aa}
		\end{array} \right).
\end{equation}
\end{linenomath*}
Note that this equation is identical to the classical equation for frequencies of male gametes in a sexually antagonistic model except that now we have to specify that only adults contribute to the gamete pool whereas in a classical population genetic model with no population structure, we do not distinguish between reproducing adults and non-reproductive juveniles. Of course equation (\ref{eq:maleGametePool_example}) starts deviating from the classical population genetics case more drastically if we include multiple reproducing adult age or stage classes (see equation (\ref{eq:maleGametePool})).  }

\subsubsection*{Population projection}
Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$) and the allele frequencies in the gamete pool, we can construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$, see Online Supplementary Material for the step by step construction of the model. The resulting matrix that projects the eco-evolutionary dynamics is:
\begin{linenomath*}
\begin{equation} \label{eq:Atilde}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
			\end{array} \right)}_{\tilde{\mbf{F}}},
\end{equation}
\end{linenomath*}
where $C$ denotes the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and  $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.

The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of extant individuals produced by selfing or outcrossing ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$). The survival matrices for individuals produced through selfing and outcrossing are,
\begin{linenomath*}
\begin{eqnarray} 
	\mcal{U}^S  = 
		\left(
			\begin{array}{ccc}
				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
			\end{array} \right),\label{eq:BlkUS}\\
				\mcal{U}^X  = 
		\left(
			\begin{array}{ccc}
				\mbf{U}^{X}_{AA} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{U}^{X}_{Aa} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{U}^{X}_{aa} \\
			\end{array} \right).\label{eq:BlkUX}
\end{eqnarray}
\end{linenomath*}
Since individuals cannot change their genotype once they are born, the survival matrices are block diagonal.
